---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/README-",
  out.width = "100%"
)
```

# glca
<!-- badges: start -->
[![CRAN RStudio mirror downloads](https://cranlogs.r-pkg.org/badges/grand-total/glca?color=blue)](https://r-pkg.org/pkg/glca)
[![CRAN RStudio mirror downloads](https://cranlogs.r-pkg.org/badges/last-month/glca?color=orange)](https://r-pkg.org/pkg/glca)
[![CRAN RStudio mirror downloads](https://cranlogs.r-pkg.org/badges/last-day/glca?color=red)](https://r-pkg.org/pkg/glca)

<!-- badges: end -->

Fits latent class analysis (LCA) including group variable and covariates. 
The group variable can be handled either by multilevel LCA described in Vermunt (2003) <DOI:10.1111/j.0081-1750.2003.t01-1-00131.x> or standard LCA at each level of group variable. 
The covariates can be incorporated in the form of logistic regression (Bandeen-Roche et al. (1997) <DOI:10.1080/01621459.1997.10473658>).

## Installation

You can install the released version of glca from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("glca")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("kim0sun/glca")
```
## Example

```{r example}
library(glca)
data("gss12")
# LCA
lca = glca(item(DEFECT, HLTH, RAPE, POOR, SINGLE, NOMORE) ~ 1,
            data = gss12, nclass = 3, verbose = FALSE)
summary(lca)
plot(lca)

# Multiple-group LCA (MGLCA)
mglca = glca(item(DEFECT, HLTH, RAPE, POOR, SINGLE, NOMORE) ~ 1,
             group = DEGREE, data = gss12, nclass = 3, verbose = FALSE)
summary(mglca)
plot(mglca)

# Multiple-group LCA with covariate(s) (MGLCR)
mglcr = glca(item(DEFECT, HLTH, RAPE, POOR, SINGLE, NOMORE) ~ SEX,
             group = DEGREE, data = gss12, nclass = 3, verbose = FALSE)
summary(mglcr)
coef(mglcr)

data("nyts18")
# Multilevel LCA (MLCA)
mlca = glca(item(ECIGT, ECIGAR, ESLT, EELCIGT, EHOOKAH) ~ 1, group = SCH_ID, 
            data = nyts18, nclass = 3, ncluster = 2, verbose = FALSE)
summary(mlca)
plot(mlca)

# MLCA with covariate(s) (MLCR)
# (HOME: level-1 covariate, School_lev: level-2 covariate)
mlcr = glca(item(ECIGT, ECIGAR, ESLT, EELCIGT, EHOOKAH) ~ SCH_LEV, group = SCH_ID,
            data = nyts18, nclass = 3, ncluster = 2, verbose = FALSE)
summary(mlcr)
coef(mlcr)
```


